Recruiting Chapter

### 2021-06-08 text/front_end_june_2021.Rmd
created revised draft of front end

---
title: "Recruiting Chapter"
author: 
  - Ozan Jaquette
  - Crystal Han
  - Irma Castaneda
bibliography: ./../other/burd_recruiting_chapter.bib
output: 
  bookdown::pdf_document2:
    toc: FALSE
csl: ./../other/apa.csl
urlcolor: blue
header-includes:
      - \usepackage{pdflscape}
      - \usepackage{geometry}
      - \usepackage{floatrow}
      - \floatsetup{capposition=top}
---

@@

1. How does the scale of visits to private high schools by public research universities compare to the scale of visits to private high schools by private universities?
1. How do the characteristics of private high schools visited by public research universities compare to the characteristics of private high schools visited by private universities?
1. To what extent do public research universities and private universities visit overlapping sets of private high schools?

@@

### Conceptualizing visits as an indicator of a relationship

@@

@@

Our broader project collected data on off-campus recruiting visits made by a convenience sample of colleges and universities during the 2017 calendar year. The data collection sample comes from three different lists of postsecondary institutions: all public research-extensive universities as defined by the 2000 Carnegie Classification (N=102); all private universities in the top 100 of U.S. News and World Report National Universities rankings (N=58); and all private colleges in the top 50 of U.S. News and World Report Liberal Arts Colleges rankings (N=47). For each of these institutions, we investigated their admissions website for pages that posted their upcoming off-campus recruiting visits. For institutions that posted such pages, we scraped the pages once per week throughout 2017. Many colleges and university only posted certain kinds of events (e.g., hotel receptions and national college fairs) but not others (e.g., day-time visits to high schools). These institutions are excluded from the analyses. Our final analysis sample consists of 17 public research universities, 13 private research universities, and 12 private liberal arts colleges. <!--[MAYBE ADD TABLE/FIGURE OR APPENDIX TABLE/FIGURE ON CHARACTERISTICS OF INSTITUTIONS IN OUR SAMPLE?]-->

@@

>>>>

---
title: "Recruiting Chapter"
author: 
  - Ozan Jaquette
  - Crystal Han
  - Irma Castaneda
bibliography: ./../other/burd_recruiting_chapter.bib
output: 
  bookdown::pdf_document2:
    toc: FALSE
csl: ./../other/apa.csl
urlcolor: blue
header-includes:
      - \usepackage{pdflscape}
      - \usepackage{geometry}
      - \usepackage{floatrow}
      - \usepackage{setspace}
      - \doublespacing
      - \floatsetup{capposition=top}
fontsize: 12pt
---

@@

1. How does the scale of visits to private high schools by public research universities compare to the scale of visits to private high schools by selective private universities?
1. How do the characteristics of private high schools visited by public research universities compare to the characteristics of private high schools visited by selective private universities?
1. To what extent do public research universities and selective private universities visit overlapping sets of private high schools?

@@

### Visits as an indicator of a social relation

@@

# Social network analysis concepts

Having conceptualized a recruiting visit as an indicator of a social relation, this section introduces basic concepts from social network analysis and motivate the research questions that guide empirical analyses.

A social network consists of a set of actors -- referred to as "vertices" -- and the connections -- referred to as "network ties" or "edges" between these actors. "One-mode" networks consist of vertices of a "type." For example, the social network of Facebook consists of users (vertices) who are connected to one another via friendship ties (edges). Similarly, the social network of Twitter consists of handles (vertices) who are connected to one another by following or being followed.

Whereas "one-mode" networks consist of vertices of the same "type" (e.g., in a publication network each vertex is an author), "two-mode" networks consist of vertices associated with one type of actor/entity having connections to vertices of another type. For example, an actor-movie network consists of actors (mode 1) who appear in movies (mode 2), and an actor shares an edge with a movie if the actor appears in the movie. The most commonly analyzed two-mode network in social networks literature is a corporate board-director network  [e.g., @RN4663]. These networks consists of directors (mode 1) and organizational boards (mode 2) and a particular director shares an edge with a particular organizational board if they sit on that board. Similarly, our recruiting visit data form a two-mode network, consisting of private high schools (mode 1) and college (mode 2). An off-campus recruiting visit from a college/university to a private high school is the "edge" that connects two particular vertices. In this network, edges can only occur between vertices of a different type; that is a college can visit a high school, but colleges cannot visit other colleges and high schools cannot receive visits from other high schools.

To reduce complexity, two-mode network graphs are often analyzed as one-mode networks [@RN2457;@RN4663]. Our two-mode college-school network can be transformed into a one-mode college network in which each vertex is a college/university and two colleges share an edge if both colleges visited at least one high school in common. The *weight* of an edge indicates the number of high schools that both colleges visited. For example, if there are 200 private high schools that received a visit from both Notre Dame and Villanova, the weight of the edge between Notre Dame and Villanova is 200.

## RQ1: Degree centrality

<!-- 
In social network analysis, *centrality* is a notion of how important or prominent or "in the middle of things"
*centrality* refers to the importance or prominence of an actor/vertex in a social network [CITE].
-->

In social network analysis, *centrality* refers to the importance or prominence or "being in the middle of things" of an actor/vertex in a social network [CITE]. Many alternative measures of centrality have been developed. All measures agree that centrality is an actor/vertex-level construct (e.g., each college in our network will have a centrality value, regardless of which centrality measure we use). However, different measures of centrality are based on different theoretical conceptualizations -- for example, being directly connected to the most actors versus being the intermediary between otherwise disconnected actors -- and are operationalized using different mathematical algorithms.

The simplest measure of centrality, *degree centrality* measures the number of edges directly connected to a vertex. For each college in our network, degree centrality is simply the number of different private high schools visited by the college. For each high school in our network, degree centrality is simply the number of colleges in our network that visited the high school. Social network theory conceptualizes degree centrality as a measure of "local" centrality, capturing the number of direct connections for each actor/vertex. For our empirical context, we conceive of degree centrality as a measure of scale of the number of private high schools a college visits and, equivalently, the number of private high schools we observe haging a social relation with each college. Given that we conceptualize visits to private high schools as a behavior consistent with the social reproduction orientation of selective private colleges/universities, we are interested in comparing how scale differs between public and private colleges/universities. This interest motivates research question 1, "how does the scale of visits to private high schools by public research universities compare to the scale of visits to private high schools by selective private universities?" We answer this research question by comparing degree centrality across universities.

<!-- 
Social network theory conceptualizes degree centrality as a measure of "local" centrality; it captures the number of direct connections but it does not capture the extent to which an actor/vertex is indirectly connected to others or whether an actor is connected to highly-connected others. 
-->

Degree is a limited measure of centrality in that only identifies direct connections. By contrast, *k-path centrality* centrality measures the extent to which an actor/vertex is indirectly connected to others, *betweenness centrality* measures the extent to which a vertex is an intermediary between otherwise disconnected groups of actors, and *eigenvector centrality* measures the extent to which a vertex is connected to highly connected others. Betweenness centrality tends to be relevant for empirical contexts (e.g., which manager gets credit for an idea) where an actor who occupies a position between groups can benefit by being a broker of information/ideas [@RN975]. By contrast, this chapter is more concerned with whether a particular college tends to have indirect connections with some colleges but not others, and what is the subset of high schools this group of colleges tends to visit. Research question 3, motivated below, explores these ideas.

## RQ2: Ego networks and homophily

Because the data structure of networks is complicated, a useful first step is to analyze "ego networks" rather than the entire social network. An "ego" is a single "focal" node (e.g., The University of Notre Dame). An ego network consists "of a focal node ('ego') and the nodes to whom ego is directly connected to (these are called 'alters') plus the ties, if any, among the alters" (p. XXX)<!-- [from http://www.analytictech.com/networks/egonet.htm#:~:text=Ego%20networks%20consist%20of%20a,form%20The%20human%20social%20network ]-->. For example, the University of Notre Dame ego network consists of all private high schools that received at least one visit from Notre Dame. Additionally, for each high school visited by Notre Dame, the ego network may include all colleges/universities in our sample that also visited the high school. This chapter analyzes recruiting visits to private high schools by 41 colleges and universities, which we can think of as 41 ego networks.

A common finding in social networks research is that actors are more likely to form connections with actors who are are similar to them [@RN1722]. Said differently, ego networks are biased towards "homophily" rather than "heterophily." Homophily is the idea that two vertices are more likely to be connected if they have similar characteristics (e.g., two people are more likely to be friends on Facebook if they have similar political ideologies). Heterophily, the opposite of homophily, is when pairs of connected vertices have different characteristics from one another.

In our analyses of recruiting visits from colleges to private high schools, we are interested in the existence of homophily versus heterophily for two reasons. First, prior scholarship suggests that private colleges and private high schools are more likely to have a relationship when they share common ideologies or view themselves as belonging to the same echelon [@RN3519; @RN4321;@RN4407]. Second, with respect to our substantive interest the behavior of public research universities, the characteristics of visited private high schools have consequences for subsequent enrollment composition of public research universities. Therefore, we are interested in whether universities tend to visit high schools that are similar to them on certain characteristics. Additionally, we are interested in whether public research universities favor high schools with certain characteristics, independent of homophily. These interests motivate research question 2, "How do the characteristics of private high schools visited by public research universities compare to the characteristics of private high schools visited by private universities?"

Decisions about which particular characteristics to analyze were driven by substantive considerations and data limitations. In particular, we analyzed geographic region, religious affiliation, academic reputation, racial composition, and enrollment size. We analyze geographic region because, for most colleges and universities, enrollment demand is likely to be stronger at high schools in the same geographic region, but nationally prestigious colleges and universities may experience high demand throughout the country. We analyze religious affiliation because private colleges are more likely to have a relationship with private high schools with a similar religious affiliation. Although public universities are formally secular, we are interested in whether particular public universities tend to target religiously affiliated private high schools. We analyze academic reputation based on the idea that more highly ranked colleges are likely to target highly ranked private high schools. Racial composition is of substantive interest because the racial composition of visited high schools has consequences for the rational composition of colleges and universities. Given our interests in social reproduction versus social mobility, we would also analyze tuition price and household income, but these data are not avalable for private high schools.

## RQ3: Network similarity and community detection

This chapter conceptualizes recruiting visits to private high schools as a behavior consistent with the social reproduction orientation of selective private colleges and universities. We are interested in how similar the recruiting networks of public research universities are to those of private colleges and universities, thereby motivating research question 3: to what extent do public research universities and private universities visit overlapping sets of private high schools? This research question is addressed, simultaneously, through the use of network graphs and by allocating each college/university to a "community" cluster.

Network graphs show vertices as nodes and edges as lines connecting two nodes. Conceptually, we think of nodes as repelling one another and edges as pulling nodes together. Therefore, nodes that share few connections are located apart and nodes that share many connections are located close together. Nodes that are very "central" -- many direct and indirect connections to other nodes in the network -- are located near the center of the graph, while nodes that are relatively disconnected are located at the periphery of the graph.  In the analyses below, we graph the one-mode college network, using relative proximity of nodes to convey whether a pair colleges have a similar network structure.

We also address research question 3 by using cluster analysis to identify colleges/universities that share common network structure. Many social network analyses utilize data-driven "community detection" algorithms to categorize vertices into a small number of groups (or cluster) based purely on patterns of network ties. Generally, groups are chosen to maximize the number of within-group ties while minimizing ties between members of different groups. We performed hierarchical cluster analysis to allocate colleges and universities into clusters based on their private high schools visits

@@

## Data

Our broader project collected data on off-campus recruiting visits made by a convenience sample of colleges and universities during the 2017 calendar year. The data collection sample comes from three different lists of postsecondary institutions: all public research-extensive universities as defined by the 2000 Carnegie Classification (N=102); all private universities in the top 100 of U.S. News and World Report National Universities rankings (N=58); and all private colleges in the top 50 of U.S. News and World Report Liberal Arts Colleges rankings (N=47). For each of these institutions, we investigated their admissions website for pages that posted their upcoming off-campus recruiting visits. For institutions that posted such pages, we scraped the pages once per week throughout 2017. Many colleges and university only posted certain kinds of events (e.g., hotel receptions and national college fairs) but not others (e.g., day-time visits to high schools). These institutions are excluded from the analyses. Our final analysis sample consists of 15 public research universities, 14 private research universities, and 12 private liberal arts colleges. <!--[MAYBE ADD TABLE/FIGURE OR APPENDIX TABLE/FIGURE ON CHARACTERISTICS OF INSTITUTIONS IN OUR SAMPLE?]-->

@@

## Methods

# Results

## Research question 1: scale of visits

Research question 1 asks, how does the scale of visits to private high schools by public research universities compare to the scale of visits to private high schools by selective private universities? This question can be answered by counting the number of visits to private high schools, which equivalent to calculating degree centrality for each university. In addition to scale, we are also interested in the relative scale of visits to private high schools. That is, the number of visits to private high schools as a proportion of all off-campus recruiting visits, or as a proportion of all recruiting visits to high schools.  Appendix Figures X and X show the number of visits by "type" (public high school, private high school, community college, other), respectively, for public research universities and selective private universities. Figures X and X show the number of visits to private high schools and public high schools, respectively, for public research universities and selective private universities, with separate bars for in-state and out-of-state visits.

INSERT FIGURE X AND FIGURE X ABOUT HERE

## Research question 2: characteristics of visited schools

TEXT

## Research question 3: overlapping network structure

TEXT
